---
title: "p8105_hw1_dx2262"
output: github_document
date: "2025-09-20"
---

```{r}
library(tidyverse)
library("moderndive")
```

# Problem 1

```{r}
# load data
data("early_january_weather")
ejw_df <- early_january_weather
```

The variables in this early_january_weather dataset are origin, year, month, day, 
hour, temp, dewp, hunid, wind_dir, wind_speed, wind_gust, precip, pressure, visib, and time_hour.
It has `r nrow(ejw_df)` rows and `r ncol(ejw_df)` columns.
The mean temperature is `r mean(pull(ejw_df, temp))`

```{r}
# create a scatterplot
ggplot(ejw_df, aes(x = time_hour, y = temp)) + geom_point(aes(colour = humid))
ggsave("ejw_plot.png")
```

The temperature data in this figure show periodic fluctuations, consistent with the daily pattern of temperature changes. In addition, the overall trend of the temperature is upward, indicating that the weather gradually warmed over these days. During a certain period, the humidity was higher, which may indicate precipitation.


# Problem 2

```{r}
df1 = 
  tibble(
    random_samp = rnorm(10),
    greater_0 = random_samp > 0,
    char_vec = sample(letters, 10, replace = TRUE),
    factor_vec = factor(sample(c("A", "B", "C"), 10, replace = TRUE))
  )

```

Try to take the mean of each variable in your dataframe:
The mean of the random sample is `r mean(pull(df1, random_samp))`, the mean of the logical vector is `r mean(pull(df1, greater_0))`.
The other two do not work because they are not numeric variables.

```{r}
# convert variables to numeric
num_logical <- as.numeric(df1$greater_0)
num_char <- as.numeric(df1$char_vec) 
num_factor <- as.numeric(df1$factor_vec) 

num_logical
num_char
num_factor
```
The logic vector is turned to numeric by converting True to 1 and False to 0. The factor vector is turned by labeling factors as numbers.
Yet we cannot convert arbitrary letters to numbers directly. 
